### 2021-06-23 thesis/tex/nsiTheory.tex
update

Compared to other fermions, the lightness of neutrino masses along with their sparse SM interactions might 
indicate that these particles provide the best starting point for us to probe new physics, in which 
exotic neutrino interactions might occur~\cite{gavela2009}. We call these interactions non-standard interactions (NSI) in order to 
distinguish them from the `standard' interactions of the Standard Model.

@@

% Moreover, it is common to subtract $\emm$ from the diagonal, which
% we are free to do since any multiple added or subtracted from diagonal does not affect the eigenvectors of the Hamiltonian and thus not the
% probabilities. So just as we subtracted away $V_{NC}$ back in Eq.~\ref{eq:components}, we can subtract $\emm$ and rotate it away.
% Our NSI potential matrix now takes the form 
% \begin{align}
%     \begin{pmatrix}
%         \epsilon_{ee} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
%         \epsilon_{e \mu} & \epsilon_{\mu\mu} & \epsilon_{\mu\tau}  \\
%         \epsilon_{e\tau} & \epsilon_{\mu\tau} & \epsilon_{\tau\tau}
%     \end{pmatrix} 
%     - \epsilon_{\mu\mu}
%     \begin{pmatrix}
%         1 & 0 & 0 \\
%         0 & 1 & 0 \\
%         0 & 0 & 1 
%     \end{pmatrix}
%      &= \begin{pmatrix}
%         \epsilon_{ee}- \epsilon_{\mu\mu} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
%         \epsilon_{e \mu} & 0 & \epsilon_{\mu\tau}  \\
%         \epsilon_{e\tau} & \epsilon_{\mu\tau} & \epsilon_{\tau\tau} - \epsilon_{\mu\mu}
%     \end{pmatrix} \nonumber \\
%     &= \begin{pmatrix}
%         \epsilon_{ee}- \epsilon_{\mu\mu} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
%         \epsilon_{e \mu} & 0 & \epsilon_{\mu\tau}  \\
%         \epsilon_{e\tau} & \epsilon_{\mu\tau} & \ep
%     \end{pmatrix}\,,
% \end{align}
% where $\ep = \ett - \emm$.

@@

Since all standard matter potentials are diagonal, 
the elements $\epsilon_{\a\b}, \a = \b$ will directly adjust the matter potential felt 
by flavor $\a$. The off-diagonal terms have a more interesting theoretical implication as the open up matter interactions across flavors. Remember, in the Standard model,
we are restriced to weak interactions that conserve lepton number. However, a off-diagonal NSI parameter allows flavor transitions during matter interactions.
Moreover, $\epsilon_{\a\b}$ modifies the $\nu_\a \to \nu_\b$ transition independently
of the mixing matrix. Thus, we can in theory boost a flavor transition which is supressed by the mixing matrix. Moreover, since we have 6 NSI parameters, and the PMNS
matrix is parametrized with only 3 parameters, we have more degrees of freedom when adjusting the probabilities using the NSI matrix. Each term in the PMNS matrix consists of at least
two of the three mixing angles, making them much more dependent on each other.

>>>>

Compared to other fermions, the lightness of neutrino masses\footnote{The extreme lightness might be explained 
by a mass generation at a higher order theory.} along with their sparse SM interactions might 
indicate that these particles provide the best starting point for us to probe new physics, in which 
exotic neutrino interactions might occur~\cite{gavela2009}. We call these interactions non-standard interactions (NSI) in order to 
distinguish them from the `standard' interactions of the Standard Model.

@@

By combining oscillation data and neutrino-nucleon scattering at the COHERENT experiment, 90 \% CL ranges of some of 
the NSI parameters were constrained in~\cite{coherent} to
\begin{align}\label{eq:coherentbounds}
    -0.090 < &\ett < 0.38 \nonumber \\
    -0.01 < &\emt < 0.009 \nonumber \\
    -0.073 < &\eem < 0.044 \nonumber \\
    -0.15 < &\eet < 0.13\,.
\end{align}

% we are free to do since any multiple added or subtracted from diagonal does not affect the eigenvectors of the Hamiltonian and thus not the
% probabilities. So jusb t as we subtracted away $V_{NC}$ back in Eq.~\ref{eq:components}, we can subtract $\emm$ and rotate it away.
% Our NSI potential matrix now takes the form 
% {µmmb m  m mb  mb  mbb  mb bmMoreover, it is common to subtract $\emm$ from the diagonal, which
% \begin{align} m mmb
%     \begin{pmatrix}
%         \epsilon_{ee} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
%         \epsilon_{e \mu} & \epsilon_{\mu\mu} & \epsilon_{\mu\tau}  \\
%         \epsilon_{e\tau} & \epsilon_{\mu\tau} & \epsilon_{\tau\tau}
%     \end{pmatrix} 
%     - \epsilon_{\mu\mu}
%     \begin{pmatrix}
%         1 & 0 & 0 \\
%         0 & 1 & 0 \\
%         0 & 0 & 1 
%     \end{pmatrix}
%      &= \begin{pmatrix}
%         \epsilon_{ee}- \epsilon_{\mu\mu} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
%         \epsilon_{e \mu} & 0 & \epsilon_{\mu\tau}  \\
%         \epsilon_{e\tau} & \epsilon_{\mu\tau} & \epsilon_{\tau\tau} - \epsilon_{\mu\mu}
%     \end{pmatrix} \nonumber \\
%     &= \begin{pmatrix}
%         \epsilon_{ee}- \epsilon_{\mu\mu} & \epsilon_{e\mu} & \epsilon_{e\tau}  \\
%         \epsilon_{e \mu} & 0 & \epsilon_{\mu\tau}  \\
%         \epsilon_{e\tau} & \epsilon_{\mu\tau} & \ep
%     \end{pmatrix}\,,
% \end{align}
% where $\ep = \ett - \emm$.

@@

Since all standard matter potentials are diagonal, 
the elements $\epsilon_{\a\b},\, \a = \b$ will directly adjust the matter potential felt 
by flavor $\a$. The off-diagonal terms have a more interesting theoretical implication as the open up matter interactions across flavors. Remember, in the Standard model,
we are restriced to weak interactions that conserve lepton number. However, a off-diagonal NSI parameter allows flavor transitions during matter interactions.
Moreover, $\epsilon_{\a\b}$ modifies the $\nu_\a \to \nu_\b$ transition independently
of the mixing matrix. Thus, we can in theory boost a flavor transition which is supressed by the mixing matrix. Moreover, since we have 6 NSI parameters, and the PMNS
matrix is parametrized with only 3 parameters, we have more degrees of freedom when adjusting the probabilities using the NSI matrix. Each term in the PMNS matrix consists of at least
two of the three mixing angles, making them much more dependent on each other.
